---
title: "Field Report"
author: "Kaycee Coleman"
date: '2022-06-13'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Orsted Fisheries Summary Field Report \textcolor{blue}{blue}

This document summarizes all field activities reported by the team on the Orsted Fisheries project. There are nine types of fieldwork components conducted under this contract: Trawl, Hook & Line, Trap, Baited Remote Underwater Video (BRUV), Telemetry, Towed Camera, Clam survey, eDNA, and Acoustic Zooplankton Fish Profiler (AZFP) Glider operations.   

Additionally, another project happening in the proposed wind energy area, [Orsted ECO-PAM](https://orsted-eco-pam-web-portal.srv.axds.co/) using passive acoustics to detect whales, is combined with our glider activity to summarize what has been done in this area. 

```{r, include=FALSE, echo=TRUE}
library(readxl)
library(dplyr)
library(knitr)
library(lubridate)
library(ggplot2)

# read in field log data
flog <- read_excel("~/Library/CloudStorage/OneDrive-SharedLibraries-RutgersUniversity/Orsted Wind Fisheries Monitoring - General/Reports and Reporting/1 Fieldwork Log.xlsx")
glog <- read_excel("~/Library/CloudStorage/OneDrive-SharedLibraries-RutgersUniversity/Orsted Wind Fisheries Monitoring - General/Reports and Reporting/1 Fieldwork Log.xlsx", sheet=2)
# flog <- read_excel("~/Downloads/1 Fieldwork Log.xlsx")
# glog <- read_excel("~/Downloads/1 Fieldwork Log.xlsx", sheet=2)

# mutate
names(flog) = gsub(" ", "", names(flog))
flog = mutate(flog, TotalTime = as.numeric(TotalTime))
names(glog) = gsub(" ", "", names(glog))

## make output tables
# field time
tout = as.data.frame(rbind(c("Total hours reported", round(sum(flog$TotalTime, na.rm=T))),
                           c("Total hours at sea reported", round(sum(flog$TotalTime[!flog$VesselUsed %in% "Truck-Land"], na.rm=T))),
                           c("Total hours of land based efforts reported", round(sum(flog$TotalTime[flog$VesselUsed %in% "Truck-Land"], na.rm=T))),
                           c("Total hours this month", round(sum(flog$TotalTime[month(flog$DateIn) %in% month(Sys.Date())])))))
names(tout) = c("Hours", "Total")
tout = tout %>% mutate(Total = as.numeric(Total))

time_sum = flog %>% 
  mutate(mo = month(DateIn),
         yr = year(DateIn),
         mo_yr = paste(yr, mo, "01", sep="-"),
         mo_yr = as.Date(mo_yr, format='%Y-%m-%d')) %>%
  group_by(mo_yr) %>% 
  summarise(hours = sum(TotalTime, na.rm=T))
p = ggplot() + 
  geom_bar(data = time_sum, aes(x = mo_yr, y = hours), stat="identity",fill="cornflowerblue") + 
  labs(x="Month", y="Total Hours of Fieldwork per Month") +
  theme_bw() +
  scale_x_date(date_labels = "%Y-%b", 
               breaks = as.Date(c("2021-12-01", "2022-01-01", "2022-02-01", "2022-03-01", "2022-04-01", "2022-05-01", "2022-06-01")))  + 
  theme(text = element_text(size = 10))

  #scale_x_continuous(breaks = round(seq(min(time_sum$mo_yr), max(time_sum$mo_yr)+5, by = 31),1)) 


# summary metrics
out = as.data.frame(rbind(c("Trips where Telemetry occurred", sum(flog$Telemetry %in% "yes")),
                          c("Trips where eDNA occurred", sum(flog$eDNA %in% "yes")),
                          c("Trips where Trapping occurred", sum(flog$Trap %in% "yes")),
                          c("Trips where Baited Remote Underwater Video occurred", sum(flog$BRUV %in% "yes")),
                          c("Trips where Towing a Camera occurred", sum(flog$TowedCam %in% "yes")),
                          c("Trips where AZFP Glider deployments/recoveries occurred", sum(flog$AZFPGlider %in% "yes")),
                          #c("Trips where ECO-PAM Glider deployments/recoveries occurred", sum(flog$ECOPAMGlider %in% "yes")),
                          c("Trips where Hook & Line occurred", sum(flog$'Hook&Line' %in% "yes")),
                          c("Trips where Trawling occurred", sum(flog$Trawl %in% "yes")),
                          c("Trips where Clam Surveys occurred", sum(flog$Clam %in% "yes"))))
names(out) = c("Metric","Total")
out = out %>% mutate(Total = as.numeric(Total)) %>% arrange(desc(Total))

# vessel use
vout = as.data.frame(rbind(c("R/V Resilience", sum(flog$VesselUsed %in% "R/V Resilience")),
                           c("R/V Rutgers", sum(flog$VesselUsed %in% "R/V Rutgers")),
                           c("F/V Dana Christine II", sum(flog$VesselUsed %in% "F/V Dana Christine II")),
                           c("R/V Petrel", sum(flog$VesselUsed %in% "R/V Petrel")),
                           c("F/V Darana R", sum(flog$VesselUsed %in% "R/V Darana R")),
                           c("F/V Joey D", sum(flog$VesselUsed %in% "R/V Joey D")),
                           c("Truck-Land", sum(flog$VesselUsed %in% "Truck-Land"))))
names(vout) = c("Vessel","Total")
vout = vout %>% mutate(Total = as.numeric(Total)) %>% arrange(desc(Total))

# health and safety, gear, and whales
hout = as.data.frame(rbind(c("Incident(s)", sum(flog$incident %in% "yes")),
                           c("Gear Interaction(s)", sum(flog$gear %in% "yes")),
                           c("Marine Mammal Observation(s)", sum(flog$marinemammals %in% "yes"))))
names(hout) = c("Event","Total")
hout = hout %>% mutate(Total = as.numeric(Total)) %>% arrange(desc(Total))

# marine mammal plot
mm = flog %>% group_by(generalMMgroup) %>% 
  summarize(n=n()) %>% 
  filter(!is.na(generalMMgroup)) %>%
  mutate(generalMMgroup = toupper(generalMMgroup))
p2 = ggplot() + geom_bar(data = mm, aes(x=generalMMgroup, y=n), stat="identity", fill="cornflowerblue") + 
  labs(x="General Marine Mammal Group",y="Number of Trips Where a Siting Occurred") + 
  theme_bw() + 
  theme(text = element_text(size = 10))

# gliders
gout = glog %>% 
  mutate(das = as.numeric(difftime(recoverydate, deploymentdate, units = "days")),
         l = paste("[",ID,"](",Links,")")) %>%
  rename("Recovery Date" = recoverydate, 
         "Deployment Date" = deploymentdate,
         "Days at Sea"=das, 
         "Data Link"=l) %>% 
  dplyr::select(-ID, -Links)

#             rd = last(DateIn),
#             'Days at Sea' = as.numeric(difftime(rd, dd, units = "days")),
# gout = flog %>% 
#   mutate(glidergroup = as.numeric(glidergroup)) %>%
#   filter(!is.na(glidergroup)) %>% 
#   group_by(glidergroup) %>%
#   mutate(gt = NA, gt = ifelse(AZFPGlider %in% "yes", "AZFP", "ECO-PAM")) %>%
#   summarize(dd = first(DateOut),
#             rd = last(DateIn),
#             'Days at Sea' = as.numeric(difftime(rd, dd, units = "days")),
#             gt = first(gt)) %>% 
#   rename('Deployment Date' = dd, 'Recovery Date' = rd, 'Glider Type' = gt) %>% # had to use non-character strings in summarize
#   arrange(desc('Deployment Date')) %>% 
#   dplyr::select(-glidergroup)
# col_order <- c("Glider Type", "Deployment Date", "Recovery Date", "Days at Sea")
# gout <- gout[, col_order]
```


```{r, include = TRUE, echo=FALSE, out.width='100%'} 
#fig.align = 'center'}
kable(tout, caption = "Time Metrics")
p
kable(out, caption = "Summary Metrics")
kable(vout, caption = "Vessel Use")
kable(hout, caption = "Health and Safety, Gear Interactions, and Marine Mammal Observations")
p2
opts <- options(knitr.kable.NA = "")
kable(gout, caption = "Glider Days at Sea", align = "lccrr")

```

